External Validation

### 2021-02-03 FEM_R/model_analysis.Rmd
Added validation visualisations to model_analysis.Rmd

```{r}
# Need a function here to handle the repetition
get_prevalence <- function(FEM) {

}

# Now read in male and female AgeUK data
male_ageUK <- read.csv('../../Validation/AgeUK-chronDisease-MALE.csv')
female_ageUK <- read.csv('../../Validation/AgeUK-chronDisease-FEMALE.csv')
# Rename columns
colnames(male_ageUK) <- c('Condition', 'Age 60-64', 'Age 65-69', 'Age 70-74', 'Age 75-79', 'Age 80-84', 'Age 85-89', 'Age 90-94', 'Age 95-99', 'Age 100+', 'Total Pop.')
colnames(female_ageUK) <- colnames(male_ageUK)

# Collect conditions into a list for extracting from FEM
# Will need to be written in FEM shorthand
keep_vars <- c('male', 'age', 'hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene')

# Now get FEM output for 2014
FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
FEM_2014 <- FEM_2014[keep_vars]
# extract males and females
mFEM_2014 <- FEM_2014 %>% filter(male == 1)
fFEM_2014 <- FEM_2014 %>% filter(male == 0)

m6064 <- mFEM_2014 %>% filter(age %in% (60:64))
f6064 <- fFEM_2014 %>% filter(age %in% (60:64))

m6064_prev <- data.frame((colSums(m6064) / nrow(m6064)) * 100)
f6064_prev <- data.frame((colSums(f6064) / nrow(f6064)) * 100)
colnames(m6064_prev) <- 'FEM_male_6064'
colnames(f6064_prev) <- 'FEM_female_6064'

prev_df <- data.frame(m6064_prev, f6064_prev)

# correct colnames in ageUK stuff
male_ageUK <- male_ageUK %>% filter(conditions == c('Hypertension', 'Stroke', 'Diabetes', 'Dementia'))

# Now calculate prevalence by age group for each chronic disease
# prevalence = total / affected
# Diabetes
male_prev <- data.frame(colSums(mFEM_2014) / nrow(mFEM_2014))
female_prev <- data.frame(colSums(fFEM_2014) / nrow(fFEM_2014))
colnames(male_prev) <- 'FEM_male'
colnames(female_prev) <- 'FEM_female'

>>>>

```{r echo = FALSE}

#AgeUK
male_au <- read.csv('../../Validation/AgeUK-chronDisease-MALE.csv')
female_au <- read.csv('../../Validation/AgeUK-chronDisease-FEMALE.csv')

# FEM 2014 detailed output
#FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
## extract males and females
#mFEM_2014 <- FEM_2014 %>% filter(male == 1)
#fFEM_2014 <- FEM_2014 %>% filter(male == 0)

# FEM 2014 summary
FEM <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')

# Keep only year 2014 from FEM summary
FEM <- filter(FEM, year == 2014)

# Specify the age groups to extract from FEM
ages <- c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p')
# Add m and f for male and female variables
m_ages <- paste0('m_', ages)
f_ages <- paste0('f_', ages)

# Collect the age suffixes into a vector
keep_vars <- c(m_ages, f_ages)
# Now set up the chronic diseases in a list so we can create a bigger list with all of them produced
diseases <- c('hearte', 'hibpe', 'stroke', 'diabe', 'lunge', 'cancre', 'demene')

# Don't like using loops but this one was just too simple to bother with figuring out the apply version
final <- c()
ticker <- 1
# Generating the column (variable) names we want to keep in a loop
for(item in diseases) {
  for(anotherItem in keep_vars) {
    final[ticker] <- paste0('p_', item, '_', anotherItem)
    ticker <- ticker + 1
  }
}

# Keep the right vars using varnames generated in loop
FEM <- FEM %>%
          select(c('year', all_of(final)))

# Rename some vars to remove 'p_' before merging
FEM <- FEM %>%
          rename_at(vars(starts_with('p_')),
                    funs(str_replace(., 'p_', ''))) %>%
          select(!year)

pivoted.FEM <- pivot_longer(data = FEM,
                            cols = hearte_m_6064:demene_f_100p, 
                            names_to = c('Condition', 'Gender', 'Age_Group'), 
                            names_sep = '_')

# Convert value from fraction to percentage
pivoted.FEM$value <- pivoted.FEM$value * 100

# Rename cols to enable reshaping
male_au <- male_au %>%
                dplyr::rename(m_6064 = X60.64, m_6569 = X65.69, m_7074 = X70.74, 
                       m_7579 = X75.79, m_8084 = X80.84, m_8589 = X85.89, 
                       m_9094 = X90.94, m_9599 = X95.99, m_100p = Over.100) %>%
                select(!Total)

female_au <- female_au %>%
                dplyr::rename(f_6064 = X60.64, f_6569 = X65.69, f_7074 = X70.74, 
                       f_7579 = X75.79, f_8084 = X80.84, f_8589 = X85.89, 
                       f_9094 = X90.94, f_9599 = X95.99, f_100p = Over.100) %>%
                select(!Total)

# Now need to reshape these datasets to long format
pivoted.male <- pivot_longer(data = male_au,
                             cols = m_6064:m_100p,
                             names_to = c('Gender', 'Age_Group'),
                             names_sep = '_')

pivoted.female <- pivot_longer(data = female_au,
                             cols = f_6064:f_100p,
                             names_to = c('Gender', 'Age_Group'),
                             names_sep = '_')

# Now row bind these frames together, combining male and female data
ageUK <-rbind(pivoted.male, pivoted.female)

# Now need to deal with the fact that ageUK data splits heart issues into 3 different conditions
# extract all heart related conditions for combining
ageUK_heart <- filter(ageUK, Condition == c('Coronary Heart Disease', 'Heart Failure', 'Atrial Fibrillation'))
# Aggregate the values by gender and age group
ageUK_heart <- aggregate(ageUK_heart$value, by = list(ageUK_heart$Gender, ageUK_heart$Age_Group), FUN = sum)
# Rename columns and add condition column back
colnames(ageUK_heart) <- c('Gender', 'Age_Group', 'value')
ageUK_heart$Condition <- 'Heart Problems'

# Now drop the heart problems from original ageUK and replace with aggregate
ageUK <- filter(ageUK, Condition != 'Coronary Heart Disease', Condition != 'Heart Failure', Condition != 'Atrial Fibrillation')
ageUK <- rbind(ageUK, ageUK_heart)

# Rename a few just for completeness
ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)

# Finally, add column to specify that this is ageUK data
ageUK$source <- "ageUK"

# REMOVE CONDITIONS FROM ageUK THAT ARE NOT IN FEM (may change in future)
#ageUK <- filter(ageUK, Condition != c('Asthma', 'Depression', 'Severe Mental Health Condition'))
ageUK <- filter(ageUK, Condition != 'Asthma', Condition != 'Depression', Condition != 'Severe Mental Health Conditions')

# Make copy
newFEM <- pivoted.FEM

# Rename all the conditions to fit ageUK
# NOTE: Lung disease is not a perfect match, as ageUK only contains info on COPD
#       Therefore only validation from this will be that FEM is higher than ageUK
newFEM$Condition <- gsub("hearte", "Heart Problems", newFEM$Condition)
newFEM$Condition <- gsub("hibpe", "Hypertension", newFEM$Condition)
newFEM$Condition <- gsub("stroke", "Stroke", newFEM$Condition)
newFEM$Condition <- gsub("diabe", "Diabetes", newFEM$Condition)
newFEM$Condition <- gsub("lunge", "Lung Disease (COPD)", newFEM$Condition)
newFEM$Condition <- gsub("cancre", "Cancer", newFEM$Condition)
newFEM$Condition <- gsub("demene", "Dementia", newFEM$Condition)

# Add column to specify this is FEM data
newFEM$source <- "FEM"

# Now row bind the datasets
combined <- rbind(newFEM, ageUK)
# Specify factor levels for age group
combined$Age_Group <- factor(combined$Age_Group, levels = c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p'))
# Sort
combined <- arrange(combined, source, Condition, Gender, Age_Group)

# Function to filter the combined dataset and visualise
visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Validation/ageUK/') {

  if(gender == 'm') {
    full_gender <- 'Male'
  }
  else if(gender == 'f') {
    full_gender <- 'Female'
  }

  # filter by gender and condition
  data <- data %>%
            filter(Gender == gender) %>%
            filter(Condition == cd)

  # now plot!
  p <- ggplot(data = data, aes(x = Age_Group, y = value)) +
          geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
          labs(title = paste(cd, full_gender, sep = ' '), x = 'Age Group', y = 'Proportion')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = paste0(cd, '_', gender, '.png'),
         plot = p,
         path = path,
         width = 7,
         height = 5)

  p
}

# Collect all conditions (and both genders) into a list, and apply the function above to that list
cd_list <- unique(combined$Condition)
gender_list <- unique(combined$Gender)

for(cd in cd_list) {
  for(gen in gender_list) {
    visAgeUKValid(combined, cd, gen)
  }
}

### 2021-02-03 FEM_R/validation.Rmd
Added validation visualisations to model_analysis.Rmd

# Next, lets try an all female and all male example

@@

>>>>

# Next, lets try an all female and all male example
comb <- combined

# extract genders
male <- filter(combined, Gender == 'm')
female <- filter(combined, Gender == 'f')

# try plotting each separately

ggplot(data = male, aes(x = Age_Group, y = value, width = 0.7)) +
  geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
  facet_wrap(~Condition) +
  labs(title = 'Male')

ggplot(data = female, aes(x = Age_Group, y = value, width = 0.7)) +
  geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
  facet_wrap(~Condition) +
  labs(title = 'Female')

@@

# Function to filter the combined dataset and visualise
visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/Validation/ageUK/') {

  if(gender == 'm') {
    full_gender <- 'Male'
  }
  else if(gender == 'f') {
    full_gender <- 'Female'
  }

  # filter by gender and condition
  data <- data %>%
            filter(Gender == gender) %>%
            filter(Condition == cd)

  # now plot!
  p <- ggplot(data = data, aes(x = Age_Group, y = value)) +
          geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
          labs(title = paste(cd, full_gender, sep = ' '), x = 'Age Group', y = 'Proportion')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = paste0(cd, '_', gender, '.png'),
         plot = p,
         path = path,
         width = 7,
         height = 5)

  p
}

visAgeUKValid(combined, 'Hypertension', 'm')

# Collect all conditions (and both genders) into a list, and apply the function above to that list
cd_list <- unique(combined$Condition)
gender_list <- unique(combined$Gender)

for(cd in cd_list) {
  for(gen in gender_list) {
    visAgeUKValid(combined, cd, gen)
  }
}
